---
title: "Mauna Loa CO~2~ Tracking: Personalize the Data"
author: "M Pearson"
date: "March 2, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
library(lubridate)
library(dplyr)

```

## Purpose

### **Problem**  
Climate change is one of the biggest challenges we face, but individuals have trouble conceptualizing or relating to the data.

### **Solution**   
Personalize the data to markers the user can set themselves, increasing their ability to relate to the data

### **Case Study**  
Personalized time period increase in CO~2~ from Mauna Loa Data


## Case Study Framework: the Data  

Since 1959 scientists have collected CO~2~ data from Mauna Loa in Hawaii - possibly *the* classic data set in Climate Science.

To allow users to explore this data we built a Shiny App that allows users to plot a post-1958 date of importance (suggestion: a birthday!) and calculates the rise in global CO2 since that date.

Tying CO~2~ rise to a personally meaningful time frame can help drive home how dramatically and rapidly global CO~2~ is changing.


## CO~2~ change this Millenia

```{r Load_Data}
#Load NOAA Data and process data frame for use
co2URL <- 'ftp://aftp.cmdl.noaa.gov/products/trends/co2/co2_mm_mlo.txt'
ml <- read.table(co2URL)
colnames(ml) <- c("Year","Month", "decimal","average","interpolated","trend","other" )

# Get most recent value and data from New Millenia, 
# get decimal value for plot
recentAvg <- ml[nrow(ml),6]
recentDecimal <- ml[nrow(ml),3]
milleniaCO2 <-   ml[which(ml$Year==2000 & ml$Month==1),6]
milleniaDecimal <-   ml[which(ml$Year==2000 & ml$Month==1),3] 

```

```{r testPlot}
tp <- plot_ly(ml) %>% add_trace(x=~decimal, y=~interpolated, type = 'scatter', mode = 'lines', name="Monthly Average", transforms = list(
           list(
                 type = 'filter',
                 target = 'y',
                 operation = '>',
                 value = 0
             )
      )) %>% add_trace(x=~decimal, y=~trend, type = 'scatter', mode = 'lines', name="Seasonally Adjusted Trend") %>%
        layout(legend = list(x = 0.05, y = .95), xaxis=list(title='Date'),yaxis=list(title='CO<sub>2</sub> [ppm]')) %>%
      add_markers(x=milleniaDecimal, y = milleniaCO2,name="January 2000 ",marker=list(symbol=~17 , size=25 , color="orange", text='input'))  %>%
      add_markers(x=recentDecimal, y = recentAvg,name="Today",marker=list(symbol=~17 , size=20 , color="red", text='today'))
tp
```
An increase of **`r recentAvg-milleniaCO2` ppm** in less than 20 years

## Resources

### **Explore the App yourself**  
Shiny App:  https://mpearson.shinyapps.io/maunaloa/

### **Github repository**  
https://github.com/mepearson/MaunaLoa

### **Data Credit**  
Data from Dr. Pieter Tans, NOAA/ESRL (www.esrl.noaa.gov/gmd/ccgg/trends/) and Dr. Ralph Keeling, Scripps Institution of Oceanography (scrippsco2.ucsd.edu/).
https://www.esrl.noaa.gov/gmd/ccgg/trends/data.html